Ornaments as indicators of social changes in northeastern Taiwan before and after European colonial period

### 2019-12-15 analysis/paper/paper.Rmd
adjust and save C14 age figure to figure file

The European presence in northern Taiwan started with the Spanish who founded Fort San Salvador at Keelung in 1626, and Fort San Domingo in 1629 at Tamsui. They sent missionaries to local indigenous settlements in this region [@Blusse2000, pp. 343] and kept records about their observations of indigenous communities. A Dominican priest in 1632 reported that the Taparri, an indigenous tribe from northern Taiwan, exchanged carnelian beads with other indigenous groups. This form of exchange was widespread and even the Spanish soldiers used carnelian beads as bargaining chips for gambling [@LiandWu2006, pp. 132-149]. The use of beads as prestige goods is further indicated by their role in bride price payments, and compenzation to resolve disputes [@LiandWu2006, pp. 132-149]. Other records mention that the female shamans in the tribe would use carnelian beads as magical items in ritual healing practices [@Borao2009, pp. 122-151]. Records of an indigenous funeral document the use of carnelian beads in ritual contexts, with more carnelian beads, pottery, and cloth placed into the graves of more influential people to indicate their family's higher status [@LiandWu2006, pp. 153]. While a full critical analysis of these historical accounts remains to be produced, we take them to minimally indicate that carnelian beads were already treated as prestige goods in Yilan before the arrival of Europeans. In 1642, the Dutch Vereenigde Oostindische Compagnie (VOC) defeated the Spanish and took over their forts in northern Taiwan. They introduced a feudal system in an attempt to control the indigenous communities by asking indigenous leaders to attend an annual ceremony for demonstrating their loyalty and paying tributes [@Andrade2007, ch. 9; @Kang2016, ch. 4]. The Dutch provided beads and other goods based on negotations with indigenous communities to secure alliances in the annual ceremony or during travelling [@Kang2016, ch. 6]. We might predict that the activities of the Dutch feudal system to build and maintain alliances resulted in an increase in the amount and diversity of ornaments in northeastern indigenous communities during this period.

@@

lines_for_periods <- 
  tibble(bx = c(324,  # start of Euro contact, 1626 AD
               182),  # start of Chinese contact, 1768 AD
          by = c(0, 0),
          byend = c(0.00175),
          bxend = bx)

ggplot() +
    geom_area(data = kw_ages_densities_dfs,
            aes(ageGrid,
                dens_to_one,
                group = date),
            fill = "grey80",
            alpha = 0.4,
            position = 'identity') +
  geom_line(data = kw_ages_densities_df,
            aes(ageGrid,
                densities)) +
  geom_segment(data = lines_for_periods,
              aes(x = bx, 
                  y = by, 
                  xend = bxend, 
                  yend = byend),
              colour = c("red", "blue")) +
  annotate("rect", xmin = 100, xmax = 600, 
                   ymin = 0, ymax = lines_for_periods$byend[1],
           fill = "grey80",
           alpha = 0.3) +
    annotate("text", x = 350, y = 0.0022,
           label = "Upper component",
           size = 2) +
    annotate("rect", xmin = 1300, xmax = 800,
                   ymin = 0, ymax = lines_for_periods$byend[1],
           fill = "grey80",
           alpha = 0.3) +
      annotate("text", x = 1050, y = 0.00220,
           label = "Lower component",
           size = 2)  +
  theme_minimal(base_size = 8) +
  xlab("calibrated years BP") +
  ylab("Density") +
  coord_fixed(ratio = 400000)

# to describe in your text, ?BchronDensity
# "We used the Bchron R package to compute a non-parametric phase model 
# on the radiocarbon ages via Gaussian Mixture density estimation (Parnell et al. 2008).
# FIgure X shows the results of this analysis... indicating that the peak period of occupation is around 600 BC"
```

>>>>

The European presence in northern Taiwan started with the Spanish who founded Fort San Salvador at Keelung in 1626, and Fort San Domingo in 1629 at Tamsui. They sent missionaries to local indigenous settlements in this region [@Blusse2000, pp. 343] and kept records about their observations of indigenous communities. A Dominican priest in 1632 reported that the Taparri, an indigenous tribe from northern Taiwan, exchanged carnelian beads with other indigenous groups. This form of exchange was widespread and even the Spanish soldiers used carnelian beads as bargaining chips for gambling [@LiandWu2006, pp. 132-149]. The use of beads as prestige goods is further indicated by their role in bride price payments, and compensation to resolve disputes [@LiandWu2006, pp. 132-149]. Other records mention that the female shamans in the tribe would use carnelian beads as magical items in ritual healing practices [@Borao2009, pp. 122-151]. Records of an indigenous funeral document the use of carnelian beads in ritual contexts, with more carnelian beads, pottery, and cloth placed into the graves of more influential people to indicate their family's higher status [@LiandWu2006, pp. 153]. While a full critical analysis of these historical accounts remains to be produced, we take them to minimally indicate that carnelian beads were already treated as prestige goods in Yilan before the arrival of Europeans. In 1642, the Dutch Vereenigde Oostindische Compagnie (VOC) defeated the Spanish and took over their forts in northern Taiwan. They introduced a feudal system in an attempt to control the indigenous communities by asking indigenous leaders to attend an annual ceremony for demonstrating their loyalty and paying tributes [@Andrade2007, ch. 9; @Kang2016, ch. 4]. The Dutch provided beads and other goods based on negotations with indigenous communities to secure alliances in the annual ceremony or during travelling [@Kang2016, ch. 6]. We might predict that the activities of the Dutch feudal system to build and maintain alliances resulted in an increase in the amount and diversity of ornaments in northeastern indigenous communities during this period.

@@

lines_for_periods <- 
  tibble(bx = c(324,  # start of Euro contact, 1626 AD
               182),  # start of Chinese contact, 1768 AD
          by = c(0, 0),
          byend = c(0.0020),
          bxend = bx)

kwl_ages_curve <-
ggplot() +
    geom_area(data = kw_ages_densities_dfs,
            aes(ageGrid,
                dens_to_one,
                group = date),
            fill = "grey80",
            alpha = 0.4,
            position = 'identity') +
    geom_line(data = kw_ages_densities_df,
            aes(ageGrid,
                densities)) +
    geom_segment(data = lines_for_periods,
              aes(x = bx, 
                  y = by, 
                  xend = bxend, 
                  yend = byend),
              colour = c("red", "blue")) +
    annotate("rect", xmin = 100, xmax = 600, 
                   ymin = 0, ymax = lines_for_periods$byend[1],
           fill = "grey80",
           alpha = 0.3) +
    annotate("text", x = 350, y = 0.0026,
           label = "Upper component",
           size = 2) +
    annotate("rect", xmin = 1300, xmax = 800,
                   ymin = 0, ymax = lines_for_periods$byend[1],
           fill = "grey80",
           alpha = 0.3) +
    annotate("text", x = 1050, y = 0.0026,
           label = "Lower component",
           size = 2)  +
  theme_minimal(base_size = 8) +
  xlab("calibrated years BP") +
  ylab("Density") +
  coord_fixed(ratio = 400000)

# see ?BchronDensity. "We used the Bchron R package to compute a non-parametric phase model 
# on the radiocarbon ages via Gaussian Mixture density estimation (Parnell et al. 2008).
# indicating that the peak period of occupation is around 600 BC"

ggsave(plot = kwl_ages_curve,
       here("analysis", "figures",
       "plot-c14-ages-distribution.jpg"),
       width = 120, 
       dpi = 300,
       height = 60,
       units = "mm")

knitr::include_graphics(here("analysis", 
                             "figures", 
                             "plot-c14-ages-distribution.jpg"))
